Mathematical Aspects of the Capital Market

### 2017-08-04 phd/Chapterintroduction.tex
Revised the PhD thesis.

\chapter{Introduction}\label{ch:intr}

\section{Regular Variation}
The concept of regular variation is defined by the following scaling
property: If a function $f(\cdot)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\]
then we say $f(x)$ is regularly varying with index $\alpha$.
Distribution functions, say $F(\cdot)$, whose survival function
$\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
observed in a variety of phenomena, for example, phase transition of
matter and black body radiation as studied in physics, neuronal
avalanches in biology, claim sizes of insurance mathematics and stock
returns in finance. The last application is indeed the focus of this
thesis.

When expanded to multiple dimensions, the scaling property of regular
variation is better described in terms of weak convergence to a
spectral measure: If a random vector $V$ satisfies
\[
\lim_{x \to \infty}
{
  \P(V/|V| \in \cdot, |V| > c x)
  \over
  \P(|V| > x)
}
\overset{w}{\to} c^{-\alpha} \mu_\alpha(\cdot)
\]
then we say the components of $V$ is jointly regularly varying with
index $\alpha$. Here $\mu_\alpha(\cdot)$ is a probability measure on
the unit sphere \cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
is jointly regularly varying with index $\alpha$, then each component
of it and each linear combination of its components are regularly
varying with the same index $\alpha$. This follows from Feller
\cite{feller}, p. 278. Cf. also Jessen and Mikosch
\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et al
\cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.

Thus estimating the tail index $\alpha$ is particularly important for
understanding the behaviour of a heavy-tailed series. The earliest
method proposed for this purpose is due to Hill \cite{hill1975simple}:
\[
\hat \alpha_H = \left[
  {1 \over k} \sum_{i=1}^k \log \left(
  X_{(i)} \over X_{(k+1)}
  \right)
  \right]^{-1}
\]
where $X_1, X_2, \dots$ is the series whose tail index is the subject
of estimation, and $X_{(i)}$ is the $i$-th upper order statistics of
this series. Several authors have contributed to showing the
weak consistency of the estimator $\hat \alpha_H$,
i.e. if $k \to \infty, k/n \to 0$ as $n \to \infty$, then
$\hat \alpha_H \overset{P}{\to} \alpha$.

@@

When considered as the multiplicative inverse of the parameter of the
{\em Generalized Extreme Value} distribution, there are other methods
in the literature for estimating the tail index, e.g. Pickand's 
estimator \cite{pickands1975statistical} $\hat \alpha_P$ and the
Deckers-Einmahl-de Haan estimator $\hat \alpha_{\text{DEH}}$
\begin{eqnarray*}
{1 \over \hat \alpha_P}
&=&
{1 \over \log 2}
\log {
  X_{(k)} - X_{(2k)}
  \over
X_{(2k)} - X_{(4k)}  
} \\
{1 \over \hat \alpha_{\text{DEH}}}
&=&
1 + {1 \over \alpha_H} + {1 \over 2} \left(
  {H \over \alpha_H^2} - 1
\right)^{-1}
\end{eqnarray*}
where it is similarly assumed $k \to \infty$ and $k/n \to 0$ as $n \to
\infty$. The Deckers-Einmahl-de Haan estimator makes use of Hill's
estimator $\hat \alpha_H$ and computes
\[
H = \left[
  {1 \over k} \sum_{i=1}^k \log \left(
    { X_{(i)} \over X_{(k+1)} }
  \right)^2
\right]^{-1}
\]
Apparently $1/H$ can be interpreted as the 2nd empirical moment of
$\log(X_{(i)}/X_{(k+1)})$ for $i \leq k$.
A major drawback of Pickand's and Deckers-Einmahl-de Haan's estimators
is that, when applied to estimating the tail index, they discard the
information that the tail index is always positive, hence resulting in
a larger confidence band compared with that obtained for Hill's
estimator. Therefore we stick to Hill's estimator in the empirical
work included in this thesis.

>>>>

\chapter{Introduction}\label{ch:intr}
The phenomena of heavy-tailedness is widely observed in many
disciplins of science, for example, phase transition of matter and
black body radiation as studied in physics, neuronal avalanches in
biology, claim sizes of insurance mathematics and stock returns in
finance. The last application is indeed the focus of this thesis. To
discuss the phenomena in precise terms, we introduce the concept of
regular variation.

\section{Regular Variation}
The concept of regular variation is defined by the following scaling
property: if a function $f(\cdot)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\quad
\forall c > 0
\]
then we say $f(x)$ is regularly varying with index $\alpha$.
% Distribution functions, say $F(\cdot)$, whose survival function
% $\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
% observed in a variety of phenomena, 
When expanded to multiple dimensions, the scaling property of regular
variation is better described in terms of vague convergence to a
spectral measure $\mu_\alpha$: if a random vector $V$ satisfies
\[
{
  \P(V/|V| \in \cdot, |V| > c x)
  \over
  \P(|V| > x)
},
\overset{v}{\to} c^{-\alpha} \mu_\alpha(\cdot)
\quad
x \to \infty, \forall c > 0
\]
then we say the $V$ is regularly varying with index $\alpha$. Here
$\mu_\alpha(\cdot)$ is a probability measure on the unit sphere
\cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
is regularly varying with index $\alpha$, then each component
of it and each linear combination of its components are regularly
varying with the same index $\alpha$. This follows from Feller
\cite{feller}, p. 278. Cf. also Jessen and Mikosch
\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et. al.
\cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.

Clearly, estimating the tail index $\alpha$ is particularly important
for understanding the behaviour of a heavy-tailed series. A standard
method proposed for this purpose is due to Hill \cite{hill1975simple}:
\[
\hat \alpha_H = \left[
  {1 \over k} \sum_{i=1}^k \log \left(
  X_{(i)} \over X_{(k+1)}
  \right)
  \right]^{-1}
\]
where $X_1, X_2, \dots$ is the series whose tail index is the subject
of estimation, and $X_{(i)}$ is the $i$-th upper order statistics of
this series. Several authors have contributed to showing the
weak consistency of the estimator $\hat \alpha_H$,
i.e. if $k \to \infty, k/n \to 0$ as $n \to \infty$, then
$\hat \alpha_H \overset{P}{\to} \alpha$.

@@

% When considered as the multiplicative inverse of the parameter of the
% {\em Generalized Extreme Value} distribution, there are other methods
% in the literature for estimating the tail index, e.g. Pickand's 
% estimator \cite{pickands1975statistical} $\hat \alpha_P$ and the
% Deckers-Einmahl-de Haan estimator $\hat \alpha_{\text{DEH}}$
% \begin{eqnarray*}
% {1 \over \hat \alpha_P}
% &=&
% {1 \over \log 2}
% \log {
%   X_{(k)} - X_{(2k)}
%   \over
% X_{(2k)} - X_{(4k)}  
% } \\
% {1 \over \hat \alpha_{\text{DEH}}}
% &=&
% 1 + {1 \over \alpha_H} + {1 \over 2} \left(
%   {H \over \alpha_H^2} - 1
% \right)^{-1}
% \end{eqnarray*}
% where it is similarly assumed $k \to \infty$ and $k/n \to 0$ as $n \to
% \infty$. The Deckers-Einmahl-de Haan estimator makes use of Hill's
% estimator $\hat \alpha_H$ and computes
% \[
% H = \left[
%   {1 \over k} \sum_{i=1}^k \log \left(
%     { X_{(i)} \over X_{(k+1)} }
%   \right)^2
% \right]^{-1}
% \]
% Apparently $1/H$ can be interpreted as the 2nd empirical moment of
% $\log(X_{(i)}/X_{(k+1)})$ for $i \leq k$.
% A major drawback of Pickand's and Deckers-Einmahl-de Haan's estimators
% is that, when applied to estimating the tail index, they discard the
% information that the tail index is always positive, hence resulting in
% a larger confidence band compared with that obtained for Hill's
% estimator. Therefore we stick to Hill's estimator in the empirical
% work included in this thesis.

### 2017-08-04 phd/MyThesis.tex
Some minor arrangements.
Revised the PhD thesis.

\begin{tabular}{ll}
Supervisor: & Prof.~Thomas Mikosch\\
& University of Copenhagen\\
& \\
Co-supervisor: & Associate Prof.~Jeffrey Collamore\\
& University of Copenhagen\\
& \\
Assessment committee: &  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
& University of Neuchatel, CH\\
& \\

& Prof.~Olivier Wintenberger\\
& University of Marie \& Pierre Curie, Paris\\
& \\

& Lecturer.~Jesper Lund Pedersen (chairman)\\
& University of Copenhagen\\
\end{tabular}

@@

\def\phdthesis{1}
\input{Chapterintroduction}
%------------------------------------------------------------------------------------
\input{ChapterTailParameters}
\input{ChapterImportanceSampling}
\input{Chapterextremes}
\input{ChapterBernoulli}

>>>>

\begin{tabular}{ll}
Supervisor: & Prof.~Thomas Mikosch\\
& University of Copenhagen\\
& \\
Co-supervisor: & Associate Prof.~Jeffrey Collamore\\
& University of Copenhagen\\
& \\

Assessment committee:  & Lecturer.~Jesper Lund Pedersen (chairman)\\
& University of Copenhagen\\
& \\

&  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
& University of Neuchatel, CH \\
& \\

& Prof.~Olivier Wintenberger\\
& University of Marie \& Pierre Curie, Paris
\end{tabular}

@@

\def\phdthesis{1}
\input{Chapterintroduction}
%------------------------------------------------------------------------------------
\input{ChapterTailParameters}
\input{ChapterImportanceSampling}
\input{ChapterBernoulli}
\input{Chapterextremes}
